### 2015-01-07 prelim/oral_prelim.tex
updated prelim slides

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Uses yearly maxima
    \item Discards many observations
    \item Models are fit using the generalized extreme value distribution with parameters $\mu, \sigma,$ and $\xi$
    \begin{align*}
      \Pr(Y < y) = \left\{  \begin{array}{ll}
        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\
        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Standardized distribution is unit Fr\'{e}chet or GEV(1, 1, 1)
    \begin{align*}
      \Pr(Z < z) = exp(-z^{-1})
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Peaks over threshold}
  \begin{itemize}  \setlength{\itemsep}{0.5em}
      \item Incorporates more data than block maxima
      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances
      \item The generalized Parety distribution has three parameters $\mu, \sigma$, and $\xi$
      \begin{align*}
        P(Y < y) = \left\{ \begin{array}{ll}
          1 - \left[1 - \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\
          1 - \exp \left\{ \frac{ y - \mu }{ \sigma} \right\} & \quad \xi = 0
        \end{array}\right.
      \end{align*}
      \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
  \end{itemize}
\end{frame}

\begin{frame}{Introduction to extremes}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012):
    \begin{itemize}
      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
      that for all sites, $\bs_i, i = 1, \ldots, d$,
      \begin{align*}
        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
      \end{align*}
      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize}
    \item Multivariate distributions:
    \begin{itemize}
      \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet

      \item The multivariate representation for the GEV is
      \begin{align*}
        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
      \end{align*}
      where
      \begin{itemize}
        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
        \item $H$ is a probability measure on $\Delta_d$
        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate analysis}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item computationally challenging to work with
      \item joint distribution only available in low dimension
    \end{itemize}
    \item Bayesian hierarchical model (Reich and Shaby, 2012)
    \item Pairwise likelihood approach (Huser and Davison, 2014)
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatiotemporal modeling for extreme values}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Model to analysis spatiotemporal extreme values
    \item Model objectives
    \begin{itemize}
      \item has marginal distribution with a flexible tail
      \item has asymptotic spatial dependence
      \item has computation on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
    \begin{align*}
      Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
      \item $\bX(\bs) \bbeta$ is the location
      \item $\bOmega = \bomega \bar{\bOmega} \bomega$ is a correlation matrix
      \item $\bomega = \text{diag} \left(\frac{1}{\sqrt{ab}}, \ldots, \frac{1}{ \sqrt{ab}}\right)$
      \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
      \item $\Sigma$ is a postive definite correlation matrix
      \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{$\chi$ statistic}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item The $\chi$ statistic is a measure of extremal dependence
   \item Specifically, we focus on $\chi(\bh)$ for the upper tail given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $\bh$.
  \end{itemize}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function
$\rho(\bh)$.
    \item Model properties:
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(\bh) = 0$ regardless of the strength of the correlation in the bulk of the distribution
    \item Tail is not flexible (Gaussian is light tailed)
    \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Model properties
    \begin{itemize}
    	\item Has flexible tail controlled by skewness $\alpha$ and degrees of freedom $2a$
    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(\bh) > 0$ (Padoan, 2011)
    	\item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item For this distribution, $\chi(\bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$
   \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
    \item We deal with this through a daily random partition (similar to Kim et al., 2005)
  \end{itemize}
\end{frame}

\begin{frame}{Random partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Daily random partition allows $z$ and $\sigma^2$ to vary by site
    \begin{align*}
      Y(\bs) = \bX(\bs) \bbeta + \lambda z(\bs) + \sigma(\bs) v(\bs)
    \end{align*}
    \item Consider a set of knots $\bw_{k} \sim$ Uniform that define a random partition
    $P_{1}, \ldots, P_{K}$ such that
    \begin{align*}
      P_{k} = \{s : k = \argmin_\ell|| \bs - \bw_{\ell}|| \}
    \end{align*}
    where $\bw = (w_1, w_2)$
    \item For $\bs \in P_{k}$
    \begin{align*}
      z(\bs) &= z_{k}\\
      \sigma^2(\bs) &= \sigma^2_{k}
    \end{align*}
    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
  \end{itemize}
\end{frame}

@@

\begin{frame}{$\chi(\bh)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.7\linewidth]{./plots/pot/chi-h.pdf}\\[-0.4in]
  \caption{$\chi$ plot for different data settings}
  \end{figure}
\end{frame}

\begin{frame}{Sample simulated datasets}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-t3.pdf}
  \caption{Gaussian and $t$ with 3 partitions}
  \end{figure}
\end{frame}

\begin{frame}{Sample simulated datasets}
  \centering
  \begin{figure}
  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-skew-t3.pdf}
  \caption{Gaussian and skew-$t$ with 3 partitions}
  \end{figure}
\end{frame}

\begin{frame}{Extension to space-time data}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We extend our previous model to be
    \begin{align*}
      Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
    \end{align*}
    where $t = 1, \ldots, T$ denotes the day of each observation.
    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
    \begin{align*}
      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\
      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
    \end{align*}
    are transformations to $\calR^2$
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 6 different data settings:
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
        \item $K=1$ partition vs $K=5$ partitions
    \end{itemize}
    \item 5 different models:
    \begin{itemize}
      \item Gaussian vs skew-$t$ marginal distribution
      \item $K = 1$ partition vs $K = 5$ partitions
    \end{itemize}
    \item Brier score used to determine model that gives best fit
  \end{itemize}
\end{frame}

@@

\begin{frame}{Brier score}
  \begin{itemize} \setlength{\itemsep}{0.5em}
  \item The Brier score for predicting exceedance of threshold $c$ is
  \begin{align*}
    [e(c) - P(c)]^2
  \end{align*}
  where
  \begin{itemize}
    \item $y$ is a test set value
    \item $e(c) = I[y > c]$
    \item $P(c)$ is the predicted probability of exceeding $c$
  \end{itemize}
  \item Relative Brier scores:
  \begin{align*}
    \text{BS}_\text{rel} = \frac{ \text{BS}_\text{method}}{ \text{BS}_\text{Gaussian}}
  \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Data analysis}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}

\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item Thresholding at $T=0, 50, 75$, and $85$ ppb
    \end{itemize}
    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
    \item Compare quantile and Brier scores using two-fold cross validation (Gneiting and Raftery, 2007)
    \item Mean function modeled as
    \begin{align*}
    	\beta_0 + \beta_1 \cdot \text{CMAQ}
    \end{align*}
   \end{itemize}
\end{frame}

@@

\begin{frame}{Discussion and future work}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Improvement of model performance when using partitioned models
    \item Thresholding makes results worse
    \begin{itemize}
      \item Possible numerical instability due to truncated normal distribution
    \end{itemize}
    \item Different ways to incorporate the temporal dependence
    \begin{itemize}
    	\item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
    \end{itemize}
    \item Different partition structure
    \begin{itemize}
      \item Distance weighting for each knot vs indicator functions
    \end{itemize}
    \item Knot selection
    \begin{itemize}
      \item Possible prior on the probability a knot is in the spatial domain
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial binary regression}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We observe $Y_i = I[Z(\bs_i) > T]$, an indicator variable that a continuous latent variable has exceeded a pre-specified threshold $T$.
    \item We model $\Pr[Y_i = 1]$
    \item Common examples:
    \begin{itemize}
      \item Logistic regression
      \begin{align*}
        \Pr[Y_i = 1] = \frac{ \exp(\bX(\bs_i) \bbeta) }{ 1 + \exp(\bX(\bs_i) \bbeta)}
      \end{align*}
      \item Probit regression
      \begin{align*}
        \Pr[Y_i = 1] = \Phi[\bX(\bs_i) \bbeta]
      \end{align*}
      where $\Phi$ is the standard normal distribution function
    \end{itemize}
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Motivation}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}
\end{frame}

\begin{frame}{Defining extremes}
\begin{columns}[c]
\column{.45 \linewidth}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Key in extreme value analysis is to define extremes
    \item Typically done in one of two ways
    \begin{itemize}
      \item Block maxima
      \begin{itemize}
        \item Red dots
      \end{itemize}
      \item Over threshold
      \begin{itemize}
        \item Values over threshold considered extreme
      \end{itemize}
    \end{itemize}
  \end{itemize}

  \column{.55\linewidth}
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/define_extreme.pdf}
    \caption{Monthly maximums recorded over a three years.}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Asymptotic result:
    \begin{itemize}
      \item Let $X_1, \ldots, X_n$ be i.i.d.
      \item Consider $M_n = \max(X_1, \ldots, X_n)$
      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
      \begin{align*}
        a_n^{-1} (M_n - b_n) \converged G(z)
      \end{align*}
      then $G(z)$ follows a generalized extreme value distribution
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized extreme value distribution has three parameters:
    \begin{itemize}
      \item $\mu \in \calR$ is a location parameter
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenge:
    \begin{itemize}
      \item Lose information by only considering maximum in a block
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized extreme value distribution
    \begin{align*}
      G(y) = \Pr(Y < y) = \left\{  \begin{array}{ll}
        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\[0.5em]
        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Standardized distribution is unit Fr\'{e}chet or GEV(1, 1, 1)
    \begin{align*}
      \Pr(Z < z) = exp(-z^{-1})
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized Pareto distribution has two parameters:
    \begin{itemize}
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenges:
    \begin{itemize}
      \item Sensitive to threshold selection
      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Peaks over threshold}
  \begin{itemize}  \setlength{\itemsep}{0.5em}
    \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
    \begin{align*}
      H(y) = P(Y < y) = \left\{ \begin{array}{ll}
        1 - \left[1 - \xi \left( \frac{ y - T }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\[0.5em]
        1 - \exp \left\{ \frac{ y - T }{ \sigma} \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Related to GEV distribution through
    \begin{align*}
      H(y) = 1 + \log[G(y)], \quad y \ge T
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Max-stable processes}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012):
    \begin{itemize}
      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
      that for all sites, $\bs_i, i = 1, \ldots, d$,
      \begin{align*}
        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
      \end{align*}
      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize}
    \item Marginally at each site, observations follow a generalized extreme value distribution
    \item Finite collection of sites:
    \begin{itemize}
      \item The multivariate representation for the GEV is
      \begin{align*}
        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
      \end{align*}
      where
      \begin{itemize}
        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
        \item $H$ is a probability measure on $\Delta_d$
        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate GEV challenges}
  \begin{itemize}
    \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
    \item Two common forms for $V(\bz)$:
    \begin{itemize}
      \item Symmetric logistic
      \begin{align*}
        V(\bz) = \left[\sum_{i = 1}^n \left( \frac{ 1 }{ z_i } \right)^{1/\alpha}\right]^\alpha
      \end{align*}
      \item Asymmetric logistic
      \begin{align*}
        V(\bz) = \sum_{l = 1}^L \left[\sum_{i = 1}^n \left(\frac{w_{il}}{z_i} \right)^{1 / \alpha_l} \right]^{\alpha_l}
      \end{align*}
      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate peaks over threshold}
  \begin{itemize}
    \item Not a lot of existing methods
    \item Often use max-stable methods due to the relationship between GEV and GPD
    \item Joint distribution function given by Falk et al. (2011)
    \begin{align*}
      H(\bz) = 1 - V(\bz)
    \end{align*}
    where $V(\bz)$ is defined as in the GEV
  \end{itemize}
\end{frame}

\begin{frame}{Extremal dependence: $\chi$ statistic}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item The $\chi$ statistic is a measure of extremal dependence in the tails
   \item Specifically, we focus on $\chi(h)$ for the upper tail given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $h$
  \end{itemize}
\end{frame}

\begin{frame}{Existing challenges}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item computationally challenging (Falk et al., 2011)
      \begin{itemize}
        \item Asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters
      \end{itemize}
      \item joint distribution only available in low dimension
    \end{itemize}
    \item Some recent approaches:
    \begin{itemize}
      \item Bayesian hierarchical model (Reich and Shaby, 2012)
      \item Pairwise likelihood approach (Huser and Davison, 2014)
    \end{itemize}
    \item Many opportunities to explore new methods
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatiotemporal modeling for extreme values}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Model to analysis spatiotemporal extreme values
    \item Model objectives:
    \begin{itemize}
      \item Has marginal distribution with a flexible tail
      \begin{itemize}
        \item allow for asymmetric distributions
        \item allow for heavy tails
      \end{itemize}
      \item Has asymptotic spatial dependence
      \item Has computation on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
    \item Model properties:
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible:
      \begin{itemize}
        \item light tailed
        \item symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
    \begin{align*}
      Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
      \item $\bX(\bs) \bbeta$ is the location
      \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
      \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
      \item $\Sigma$ is a postive definite correlation matrix
      \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{$\chi(h)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.65\linewidth]{./plots/pot/chi-h-t_and_gaus.pdf}\\[-0.2in]
  \caption{$\chi$ plot for symmetric $t$, skew-$t$, and Gaussian}
  \end{figure}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Model properties
    \begin{itemize}
    	\item Has flexible tail:
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
    	\item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
    \begin{itemize}
      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
      \item We deal with this through a daily random partition (similar to Kim et al., 2005)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Random partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Daily random partition allows $z$ and $\sigma^2$ to vary by site
    \begin{align*}
      Y(\bs) = \bX(\bs) \bbeta + \lambda z(\bs) + \sigma(\bs) v(\bs)
    \end{align*}
    \item Consider a set of knots $\bw_{k} \sim$ Uniform that define a random partition
    $P_{1}, \ldots, P_{K}$ such that
    \begin{align*}
      P_{k} = \{s : k = \argmin_\ell|| \bs - \bw_{\ell}|| \}
    \end{align*}
    where $\bw = (w_1, w_2)$
    \item For $\bs \in P_{k}$
    \begin{align*}
      z(\bs) &= z_{k}\\
      \sigma^2(\bs) &= \sigma^2_{k}
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Random partition}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
    \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
  \end{itemize}
\end{frame}

@@

\begin{frame}{$\chi(h)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.7\linewidth]{./plots/pot/chi-h.pdf}\\[-0.4in]
  \caption{$\chi$ plot for different data settings}
  \end{figure}
\end{frame}

\begin{frame}{Random partition skew-$t$ model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
    \begin{itemize}
      \item Marginal distribution with flexible tails
      \begin{itemize}
        \item $\lambda$ term allows for asymmetry
        \item d.f. controls heavy vs light tails
      \end{itemize}
      \item Asymptotic spatial dependence for that decays with distance between sites via partition
      \item Computation is on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Temporal dependence}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Temporal dependence should be accounted for when using daily data
    \item For multiple days of observations the model becomes
      \begin{align*}
        Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
      \end{align*}
      where $t = 1, \ldots, T$ denotes the day of each observation.
    \item Different ways to incorporate the temporal dependence
    \begin{itemize}
      \item Time series on $\bw_t$, $z_t(\bs)$, and $\sigma_t(\bs)$
      \item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
    \end{itemize}
    \item Our model incorporates the time series because the $z_t(\bs)$ and $\sigma_t(\bs)$ terms dictate the tail behavior.
  \end{itemize}
\end{frame}

% \begin{frame}{Sample simulated datasets}
%   \centering
%   \begin{figure}
%   \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-t3.pdf}
%   \caption{Gaussian and $t$ with 3 partitions}
%   \end{figure}
% \end{frame}

% \begin{frame}{Sample simulated datasets}
%   \centering
%   \begin{figure}
%   \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-skew-t3.pdf}
%   \caption{Gaussian and skew-$t$ with 3 partitions}
%   \end{figure}
% \end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 6 different data settings:
    \begin{itemize}
      \item Gaussian, $K = 1$ partition
      \item $T$, $K = 1$ partition
      \item $T$, $K = 5$ partitions
      \item Skew-$t$, $K = 1$ partition
      \item Skew-$t$, $K = 5$ partitions
      \item Max-stable
      \begin{itemize}
        \item Marginally: GEV$(\mu=1, \sigma=1, \xi=0.2)$
        \item Dependence function: asymmetric logistic with $\alpha = 0.5$
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 5 different models:
    \begin{itemize}
      \item Gaussian
      \item Skew-$t$ with $K = 1$ partition, no thresholding
      \item Skew-$t$ with $K = 1$ partition, thresholding at $q(0.80)$
      \item Skew-$t$ with $K = 5$ partitions, no thresholding
      \item Skew-$t$ with $K = 5$ partitions, thresholding at $q(0.80)$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Brier score}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Brier score used to determine model that gives best fit (Gneiting and Raftery, 2007)
    \item The Brier score for predicting exceedance of threshold $c$ is
    \begin{align*}
      [e(c) - P(c)]^2
    \end{align*}
    where
    \begin{itemize}
      \item $y$ is a test set value
      \item $e(c) = I[y > c]$
      \item $P(c)$ is the predicted probability of exceeding $c$
    \end{itemize}
    \item Relative Brier scores:
    \begin{align*}
      \text{BS}_\text{rel} = \frac{ \text{BS}_\text{method}}{ \text{BS}_\text{Gaussian}}
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Data analysis}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item Thresholding at
      \begin{itemize}
         \item $T = 0$
         \item $T = 50$ppb, $q(0.48)$
         \item $T = 75$ppb, $q(0.92)$
         \item $T = 85$ppb, $q(0.97)$
      \end{itemize}
    \end{itemize}
    \item Compare Brier scores using two-fold cross validation
  \end{itemize}
\end{frame}

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
    \item Covariate data from the Environmental Protection Agency's Community Multiscale Air Quality (CMAQ) system.
    \item Mean function modeled as
    \begin{align*}
    	\bX_t(\bs) \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t (\bs)
    \end{align*}
   \end{itemize}
\end{frame}

@@

\begin{frame}{Discussion and future work}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Improvement of model performance when using partitioned models
    \item Thresholding makes results worse
    \begin{itemize}
      \item Possible numerical instability due to truncated normal distribution
    \end{itemize}
    \item Different partition structure
    \begin{itemize}
      \item Distance weighting for each knot vs indicator functions
    \end{itemize}
    \item Knot selection
    \begin{itemize}
      \item Possible prior on the probability a knot is in the spatial domain
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Future work: temporal dependence}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We extend our previous model to be
    \begin{align*}
      Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
    \end{align*}
    where $t = 1, \ldots, T$ denotes the day of each observation.
    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
    \begin{align*}
      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\
      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
    \end{align*}
    are transformations to $\calR^2$
  \end{itemize}
\end{frame}
